An interactive environment for mapping computational structures to FPGAs

### 2016-07-07 paper/paper.tex
Paar intro-aanpassinkjes, CSDF ref toegevoegd

  FPGAs have an enormous potential in terms of computational power.
  However, they still remain hard to program even though a lot of work has been performed on translating imperative languages to hardware.
  By using a sequential language to describe parallel hardware a structural mismatch emerges.
  A lot of analysis is therefore required to determine which computations are independent and can therefore be performed in parallel.

  In this work-in-progress paper, we propose a more structural approach to computations.
  As input, a mathematical description of the application is given which is transformed into a CSDF graph.
  Transformations to enhance performance are then applied to this graph.
  A cycle accurate simulation gives insight into where the bottlenecks of the application are.
  Finally, when the performance of the graph is sufficient, the graph is translated to hardware.

@@

  Computational structures are borrowed from functional programming languages so called higher-order functions.
  These functions are very suitable to express the structure in computations and therefore a good source of parallelism~\cite{Wester15}.
  These computational structures are represented as a specific kind of dataflow graph, a cyclo static synchronous (CSDF) dataflow graph.
  Transformations are then applied to this graph to optimize performance~\cite{deGroote16}.
  During the final step, the graph is mapped to hardware by translating each node to a combinational component and replacing the edges by a FIFO.

@@

  The CSDF graph representation of this specification is depicted in Figure~\ref{fig:csdf-dotproduct}.
  Each node in the graph represents a function, where function is to be interpreted in the widest sense: we regard values simply as functions that take no argument.
  The node corresponding to the addition has a self-loop, which indicates that it is involved in a \emph{stateful} computation - the fold.
  Note that neither of the two higher-order functions that appear in the specification (fold and zipWith) are represented by nodes in the CSDF graph.
  Higher-order functions do not map to individual nodes, but rather to \emph{subgraphs}.
  In general, the CSDF graphs derived from functional specifications are \emph{hierarchical}.
  That is, one may collapse a subgraph into a single node, which makes the higher-order functions explicit.
  \begin{figure}[h!]
    \centering
    \input{csdf-dp}
    \caption{Cyclo-static dataflow graph representation of dot-product}
    \label{fig:csdf-dotproduct}
  \end{figure}

@@

  Currently, the hardware generation backend only supports HSDF graphs, i.e., the most simple type of SDF graph.
  To add support for complete SDF and CSDF graphs, the edges have to support storing a variable amount of tokens each clock cycle.
  This is in terms of hardware much more complicated.

>>>>

  FPGAs have an enormous potential in terms of computational power.
  However, they still remain hard to program even though a lot of work has been performed on translating imperative languages to hardware.
  Imperative languages have evolved around \emph{sequential} machines.
  Using such a language to describe parallel hardware therefore causes a mismatch in \emph{structue}.
  Rather than resolving this mismatch using complex dependency analyses, we propse, in this work-in-progress-paper, a \emph{structural} approach to computations.
  In our approach, we take as input a mathematical\todo{Functional, declarative?} description of the application, and transform it into a \emph{cyclo-static dataflow} (CSDF) graph~\cite{bilsen1996}.
  Transformations to enhance performance are then applied to this graph.
  A cycle accurate simulation gives insight into where the bottlenecks of the application are.
  Finally, when the performance of the graph is sufficient, the graph is translated to hardware.

@@

  Computational structures are borrowed from functional programming languages so called higher-order functions.
  These functions are very suitable to express the structure in computations and therefore a good source of parallelism~\cite{Wester15}.
  These computational structures are represented as a specific kind of dataflow graph, a cyclo-static dataflow (CSDF) graph.
  Transformations are then applied to this graph to optimize performance~\cite{deGroote16}.
  During the final step, the graph is mapped to hardware by translating each node to a combinational component and replacing the edges by a FIFO.

@@

  The CSDF graph representation of this specification is depicted in Figure~\ref{fig:csdf-dotproduct}.
  Each node in the graph represents a function, where function is to be interpreted in the widest sense: we regard values simply as functions that take no argument.
  The node corresponding to the addition has a self-loop, which indicates that it is involved in a \emph{stateful} computation - the fold.
  Note that neither of the two higher-order functions that appear in the specification (fold and zipWith) are represented by nodes in the CSDF graph.
  Higher-order functions do not map to individual nodes, but rather to \emph{subgraphs}.
  In general, the CSDF graphs derived from functional specifications are \emph{hierarchical}.
  That is, one may collapse a subgraph into a single node, which makes the higher-order functions explicit.
  \begin{figure}[h!]
    \centering
    \input{csdf-dp}
    \caption{Cyclo-static dataflow graph representation of dot-product.}
    \label{fig:csdf-dotproduct}
  \end{figure}

@@

  Currently, the hardware generation backend only supports \emph{homogeneous} SDF (HSDF) graphs, which is the simplest type of SDF graph.
  To add support for complete CSDF graphs, the edges have to support storing a variable amount of tokens each clock cycle.
  This is in terms of hardware much more complicated.
